---
title: "P2P Data Wrangling Demo using Gil's Israel Quadrat Dataset"
author: "Ben Best"
date: "8/3/2018"
output:
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T, eval = T, message = F, warning = F)
```

## Goals

Procedures:

1. Read in Excel data
1. Wrangle tables
1. Plot map
1. Plot timeseries
1. Push to website

## Read Excel Spreadsheet

```{r, eval=T, message=F}
# load libraries
library(tidyverse)  # tibble, dplyr, tidyr, ggplot2, ...
library(readxl)     # import xls
library(here)       # look for *.Rproj

xlsx <- here::here("data/raw/Israel sites fall 2015-16.xlsx")

d <- readxl::read_excel(xlsx)
d %>% head(100) %>% DT::datatable()
```

## Species

```{r}
spp <- read_excel(xlsx, range="A1:DF636") %>%  
  slice(c(623:635)) %>%
  gather(key = header, value = val, -Unique_Code) %>%
  spread(Unique_Code, val) %>%
  select(
    spp_id=Index, header, ID_Aphia, ID_NCBI, Method,
    Kingdom, Phylum, Subphylum, Class, Subclass, Order, Family, Genus,
    Author)
spp %>% head(100) %>% DT::datatable()
```

### Match Species to WoRMS ID

- [rOpenSci | taxize tutorial](https://ropensci.org/tutorials/taxize_tutorial/)

```{r}
library(taxize)
library(fs)

View(spp)
taxa_csv <- here("data/clean/taxa.csv")

w <- get_wormsid(spp$spp_id, ask = FALSE)

w
```

```{r}
# get_wormsid() is time consuming, so cache output
if (!file_exists(taxa_csv)){

  taxa <- spp %>%
    mutate(
      worms_id  = get_wormsid(spp_id, ask = FALSE))  
  
  write_csv(taxa, taxa_csv)
} else {
  taxa <- read_csv(taxa_csv, col_type = cols())
}



# TODO: answer prompts (ask = TRUE) to resolve multiple options
taxa_tomatch <- taxa %>% 
  filter(is.na(worms_id)) %>%
  mutate(
      worms_id  = get_wormsid(GenusSpecies, ask = TRUE))

get_worms_class <- function(worms_id){
  message(glue("worms_id: {worms_id}"))
  if (is.na(worms_id)) return(NA)
  classification(worms_id, db="worms")[[1]] %>% tail(1)
}

taxa %>%
  mutate(
    worms_class = map_df(worms_id, get_worms_class))

classification(taxa$wormsid[1], db="worms")[[1]] %>% tail(1)

#taxa$wormsid <- get_wormsid(taxa$GenusSpecies)
#taxa$wormsid <- get_wormsid(taxa$query)
#taxa$wormsid <- get_wormsid(taxa$query, ask = FALSE)

#table(!is.na(taxa$wormsid))
taxa %>%
  arrange(!is.na(wormsid), GenusSpecies) %>%
  datatable()
```

- out: [data/taxa.csv](https://github.com/bbest/ioos-bio-tidyr/blob/gh-pages/data/taxa.csv)


## Observations

```{r}
obs <- read_excel(xlsx, range="A1:DF616") %>%
  gather(key = spp_id, value = n, -Unique_Code) %>%
  rename(
    qrt_id = Unique_Code) %>%
  filter(
    !is.na(n), 
    n > 0)
obs %>% head(100) %>% DT::datatable()
```

## Summary of observations by species

```{r}
obs_sum <- obs %>%
  group_by(spp_id) %>%
  summarize(
    n_max = max(n),
    n_avg = mean(n),
    n_sum = sum(n),
    n_sd  = sd(n),
    n_obs = n())
obs_sum %>% head(100) %>% DT::datatable()
```

### Questions

Try creating code to answer the following questions:

1. What's the most abundant species, ie total individuals across all observations?
1. What's the most common species, ie occurs most frequently?

## Quadrats

```{r}
qrt <- read_excel(xlsx, range="A1:EC616") %>%
  select(
    qrt_id = Unique_Code, 
    Year, Season, site_id = Site_name, Trans_Type, Qrt_serial, 
    starts_with("Parameters"))
qrt %>% head(100) %>% DT::datatable()
```

### Sites

```{r}
library(ggmap)
library(stringr)
library(glue)

sites <- tibble(
    site_id=unique(qrt$site_id)) %>%
  mutate(
    loc_id = str_replace_all(site_id, "[0-9]", "") %>%
      str_replace_all("-", " ") %>% str_trim(),
    loc_isr = glue("{loc_id}, Israel") %>% as.character())
sites %>% head(100) %>% DT::datatable()
```

#### Locations

```{r}
locs <- sites %>%
  distinct(loc_id, loc_isr) %>%
  mutate(
    lon = NA)

while (nrow(filter(locs, is.na(lon))) > 0){
  locs_na <- locs %>%
    filter(is.na(lon)) %>%
    select(loc_id, loc_isr)
  
  locs <- locs %>%
    filter(!is.na(lon)) %>%
    bind_rows(
      locs_na %>%
        bind_cols(
          geocode(locs_na$loc_isr, output="more", source="google")))
  
  Sys.sleep(1)
}
locs %>% head(100) %>% DT::datatable()
```

#### Map of locations

```{r}
library(leaflet)

leaflet(locs) %>%
  addProviderTiles(providers$Esri.OceanBasemap) %>%
  addMarkers(lng = ~lon, lat = ~lat, label = ~loc_id)
```

## Relationships between tables

```{r}
library(datamodelr) # devtools::install_github("bergant/datamodelr")

dm <- dm_read_yaml(here("data/Israel_datamodelr.yml"))
graph <- dm_create_graph( 
  dm, 
  graph_attrs = "rankdir = RL", 
  edge_attrs = "dir = both, arrowtail = crow, arrowhead = tee", 
  node_attrs = "fontname = 'Arial'")
dm_render_graph(graph)
```

## Write to CSV

```{r}
tbl_to_csv <- function(tbl, dir = here("data/clean")){
  get(tbl) %>%
    write_csv(glue("{dir}/{tbl}.csv"))
}

sapply(c("spp", "obs", "obs_sum", "qrt", "sites", "locs"), tbl_to_csv)
```



## Wrangle to Darwin Core

TODO

## Basic Plots

TODO

### Boxplot between years & species

TODO



